---
title: "Lab 5: Data Wrangling"
author: "Atibha Sheoran"
date: "2/19/2020"
output: html_document
---

```{r}
setwd("~/COLLEGE BABY!!!/Senior year/S2/Bio 497G/Labs/Data")
```

```{r eval= FALSE}
library(tidyverse)
download.file(url="https://ndownloader.figshare.com/files/2292169", destfile = "~/COLLEGE BABY!!!/Senior year/S2/Bio 497G/Labs/Data/portal_data_joined.csv")
surveys <- read_csv("~/COLLEGE BABY!!!/Senior year/S2/Bio 497G/Labs/Data/portal_data_joined.csv")
```

```{r eval= FALSE}
str(surveys)
View(surveys)
```

**SELECT**
The first word in the parenthesis indicates what argument to use and the next words indicate what columns to keep
```{r}
a <- select(surveys, plot_id, species_id, weight)
a
```
The first word in the parenthesis indicates what argument to use and the next words indicate what columns NOT to keep 
```{r}
select(surveys, -record_id, -species_id)
```
**FILTER**
To choose rows based on specific criteria
```{r}
filter(surveys, year == 1995)
```
To create a temporary data frame...
```{r}
surveys2 <- filter(surveys, weight < 5)
surveys_sml <- select(surveys2, species_id, sex, weight)
```
**NEST FUNCTIONS:** one function inside of another one
```{r}
surveys_sml <- select(filter(surveys, weight < 5), species_id, sex, weight)
```
**PIPES:** take output of one function and send it directly to the next
```{r}
surveys %>% filter(weight < 5) %>% select(species_id, sex, weight)
```
To create a final data frame
```{r}
surveys_sml <- surveys %>% filter(weight < 5) %>% select(species_id, sex, weight)
surveys_sml
```

**MUTATE**
```{r}
surveys %>% mutate(weight_kg = weight / 1000)
```
```{r}
surveys %>% mutate(weight_kg = weight / 1000, weight_lb = weight_kg * 2.2)
```
**Head:** Used to cut off the table at a few rows
```{r}
surveys %>% mutate(weight_kg = weight / 1000) %>% head()
```
Using **filter** to remove N/A data _filter(!is.na(###))_
```{r}
surveys %>% filter(!is.na(weight)) %>% mutate(weight_kg = weight / 1000) %>% head()
```
**SPLIT-APPLY-COMBINE:** split data into groups, apply analysis to each group, combine the results
```{r}
surveys %>% group_by(sex) %>% summarize(mean_weight = mean(weight, na.rm = TRUE))
```
Grouping by multiple columns
```{r}
surveys %>% group_by(sex, species_id) %>% summarize(mean_weight = mean(weight, na.rm = TRUE))
```
```{r}
surveys %>% filter(!is.na(weight)) %>% group_by(sex, species_id) %>% summarize(mean_weight = mean(weight))
```
**PRINT FUNCTION:** Using _(n = ##)_ to display a certain number of rows
```{r}
surveys %>% filter(!is.na(weight)) %>% group_by(sex, species_id) %>% summarize(mean_weight = mean(weight)) %>% print(n = 15)
```
```{r}
surveys %>% filter(!is.na(weight)) %>% group_by(sex, species_id) %>% summarize(mean_weight = mean(weight), min_weight = min(weight))
```
```{r}
surveys %>% filter(!is.na(weight)) %>% group_by(sex, species_id) %>% summarize(mean_weight = mean(weight),min_weight = min(weight)) %>% arrange(min_weight)
```
```{r}
surveys %>% filter(!is.na(weight)) %>% group_by(sex, species_id) %>% summarize(mean_weight = mean(weight), min_weight = min(weight)) %>% arrange(desc(mean_weight))
```
**COUNTING**
```{r}
surveys %>% count(sex)
```
```{r}
surveys %>% group_by(sex) %>% summarize(count = n())
```
```{r}
surveys %>% count(sex, sort = TRUE)
```
```{r}
surveys %>% count(sex, species)
```
```{r}
surveys %>% count(sex, species) %>% arrange(species, desc(n))
```